Title here

### 2016-11-23 portfolio/text/ms.Rmd
Some quick edits to intro

Managing natural resources and sustaining the livelihoods of those that depend on their extraction is difficult. These professions tend to have higher year-to-year variability compared to all others [@mishra2002; @sethi2012b]. In addition to economic risk, such as changes in prices, demand, or business cycles, individuals dependent on natural resources are also subject to biological and environmental variability. Examples include drought and flooding being a major source of risk to agricultural food security and farmers' incomes [@gregory2005], frequent catastrophic disease outbreaks or wildfires having negative impact on the logging industry [@kirilenko2007], or fluctuating prices [include mineral/gas example – Eric can’t find a good one].

Individuals who rely on natural resources for income develop strategies to reduce this variability in year to year earnings. For example, farmers may rotate their crops to buffer short-term climate variation and long-term environmental trends [@zentner1988]. Similarly, loggers maintain a diverse mix of tree species, age structure, and harvest in geographical areas to buffer against disease outbreak, wildfires, and price variation REF.

In many of these socio-ecological systems, otherwise well-intentioned regulations can limit the basket of actions available to individuals to reduce income variability, or incentivize against diversification. In other words, while diversification is generally seen as reducing variability, many individuals don’t. For example, in terrestrial agriculture, crop subsidies in the US may incentivize some farms to specialize [e.g. dairy, grains, soybeans;  @mishra2002]. In other instances, external market forces (prices, demand) may discourage the adoption of diverse strategies and help promote specialization [@bradshaw2004]. In fisheries, a series of sequential changes around the world have generally limited participation and led to increased specialization [@hilborn2003a].

Commercial fishers present an interesting example of diversification, perhaps because of the extreme risk that some individuals experience, particularly relative to terrestrial agriculture (@sethi2012b). First, the majority of insurance programs available to terrestrial farmers are not available to commercial fishermen (e.g. crop insurance programs; @kasperski2013). Second, as both fish and fishers are spatially mobile, equating fishers to farmers would be possible if the farmers’ field became an amorphous blob that changed shape and moved unpredictably day to day. Fishermen still have the ability to lower risk, however. While the days of open, unregulated fisheries in most developed countries have passed (fishers must own a license to participate in virtually all fisheries), fishers can (1) fish multiple species as permitted by a license, (2) own multiple licenses, or (3) diversify spatially and fish in multiple regions.

Despite these ways fishers can diversify, the effect of adopting various diversification strategies for improving revenue and reducing revenue variability remains uncertain. Furthermore, theory and empirical studies both suggest there can be substantial benefits to specialization (REFS). The majority of papers that examine effects of diversification on revenue do so at aggregate levels (households, counties, communities), rather than focus on individuals. Because of this, key uncertainties remain about the fundamental relationship between diversity, fishing revenue, and fishing revenue variability at the level of the individual.

One of the most well-studied examples of diversification are the state and federal fisheries in Alaska. These provide a unique opportunity to understand the role of individuals' strategies in determining income and income variability for those whose livelihood depends on the extraction of natural resources. Fisheries in Alaska are among the highest-valued fisheries in the world with 6 billion landed pounds valued at nearly 2 billion USD annually [@nmfs2016]. The industry supports a range of fishing fleets ranging from small, open, single-operator vessels to multi-million-dollar state-of-the-art ships with large crews. Like much of the rest of the world [@hilborn2003a], the days of open, unregulated fisheries in most developed countries have passed --- fishers must own a license to participate in virtually all fisheries. And while such regulations have proven helpful in reducing or eliminating overfishing, they do limit the available strategies to reduce income variability.

Using a unique dataset collected over 30 years (1985--2014), collected on individuals, rather than fleets or vessels, we are interesting in how revenue variability of individuals changes through time, and is affected by diversification. The benefit of using individuals rather than vessels is that we are able to track fishes that switch fishery sectors over time, or alter strategies and utilize multiple vessels. We developing and apply a novel statistical approach for understanding the effect of diversity on both revenue and risk that builds on results from the vessel [@kasperski2013], fleet [@sethi2012b], and community level variability [@sethi2014]. To our knowledge, this is the first application of this type of model in any natural resource field that jointly explores the effect of diversity on revenue and risk at the individual level and the first considers the multiple axes of diversity simultaneously.

>>>>

Managing natural resources and sustaining the livelihoods of those that depend on their extraction is difficult. These professions tend to have high year-to-year variability in income [@mishra2002; @sethi2012b]. In addition to economic risk, such as changes in prices or demand, individuals dependent on natural resources are also subject to biological and environmental variability. For example, drought and flooding are a major source of risk to agricultural food security and farmers' incomes [@gregory2005], and catastrophic disease outbreaks or wildfires increase risk for the logging industry [@kirilenko2007].

Individuals who rely on natural resources for income develop strategies to reduce this variability in income. For example, farmers may rotate their crops to buffer short-term climate variation and long-term environmental trends [@zentner1988]. Similarly, loggers maintain a diverse mix of tree species, age structure, and harvest in geographical areas to buffer against disease outbreak, wildfires, and price variation REF.

In many of these socio-ecological systems, otherwise well-intentioned regulations can limit the actions available to individuals to reduce income variability, or incentivize against diversification. For example, in terrestrial agriculture, crop subsidies in the US may incentivize some farms to specialize [e.g. dairy, grains, soybeans;  @mishra2002]. In other instances, external market forces (prices, demand) may discourage the adoption of diverse strategies and help promote specialization [@bradshaw2004]. In fisheries, a series of sequential changes around the world have generally limited participation and led to increased specialization [@hilborn2003a].

Commercial fishers present an interesting example of diversification, given the extreme risk that some individuals experience, particularly relative to terrestrial agriculture (@sethi2012b). First, the majority of insurance programs available to terrestrial farmers are not available to commercial fishermen [e.g. crop insurance programs; @kasperski2013]. Second, as both fish and fishers are spatially mobile, equating fishers to farmers would be possible if the farmers’ field became an amorphous blob that changed shape and moved unpredictably day to day. Fishermen still have the ability to lower risk, however. While the days of open, unregulated fisheries in most developed countries have passed (fishers must own a license to participate in virtually all fisheries), fishers can fish multiple species as permitted by a license, own multiple licenses, or diversify spatially and fish in multiple regions.

Despite these ways fishers can diversify, the effect of adopting various diversification strategies for improving revenue and reducing revenue variability remains uncertain. Furthermore, theory and empirical studies both suggest there can be substantial benefits to specialization (REFS). The majority of papers that examine effects of diversification on revenue do so at aggregate levels (e.g. households, counties, communities; REFS), rather than focus on individuals. Because of this, key uncertainties remain about the fundamental relationship between diversity, fishing revenue, and fishing revenue variability at the level of the individual.

One of the most well-studied examples of diversification are the state and federal fisheries in Alaska. These provide a unique opportunity to understand the role of individuals' strategies in determining income and income variability for those whose livelihood depends on the extraction of natural resources. Fisheries in Alaska are among the highest-valued fisheries in the world with 6 billion landed pounds [TODO kg] valued at nearly 2 billion USD annually [@nmfs2016]. The industry supports a range of fishing fleets ranging from small, open, single-operator vessels to multi-million-dollar state-of-the-art ships with large crews. Like much of the rest of the world [@hilborn2003a], the days of open, unregulated fisheries in most developed countries have passed --- fishers must own a license to participate in virtually all fisheries. And while such regulations have proven helpful in reducing or eliminating overfishing, they do limit the available strategies to reduce income variability.

Using a unique dataset collected over 30 years (1985--2014) collected on individuals, rather than fleets or vessels, we examine how revenue variability of individuals changes through time, and is affected by diversification. The benefit of using individuals rather than vessels is that we are able to track fishers that switch fishery sectors over time, or alter strategies and use multiple vessels. We develop and apply a novel statistical approach for understanding the effect of diversity on both revenue and risk that builds on results from the vessel [@kasperski2013], fleet [@sethi2012b], and community level variability [@sethi2014]. To our knowledge, this is the first application of this type of model in any natural resource field that jointly explores the effect of diversity on revenue and risk at the individual level and the first considers the multiple axes of diversity simultaneously.
